---
title: ""
output: learnr::tutorial
runtime: shiny_prerendered
---


# BTracT Dashboard

<br><br>

```{r, echo=TRUE,include=F, results='hide'}
library(dplyr)
library(tidyr)
library(DT)
library(data.table)
library(ggplot2)
library(highcharter)
banana = fread("bananadata.csv")
```

## Graphs

The graphs in this dashboard will provide more information when you mouse-over elements of the graph.
Most of the graphs are interactive. By using global filters or clicking bars of these graphs expands to specific details.

The image of the graph may be downloaded in PNG or other formats as provided in the graph menu. 
To save the plot, right click on it and then select <b> save picture as </b>. In the resulting dialog box, select the location to save the image and type in the image name in the <b> File name box</b>.


```{r, echo=FALSE}

datevalue = banana %>% dplyr::select(contains("Date")) %>%
  gather(d, date, contains("Date"), na.rm=T) %>%
  dplyr::select(date) %>% unique()
  
datemin = min(lubridate::ymd(datevalue$date))
datemax = max(lubridate::ymd(datevalue$date))
column(3, selectInput("site1", "Select Site:", c("All","Arusha","Sendusu"), selected = "All", multiple = F))
column(3, dateRangeInput("dateRange1", "Select Date Range", min = datemin, max=datemax, start=datemin, end=datemax))


result <- banana %>%
      dplyr::select(Location,Crossnumber,Mother, Father,`First Pollination Date`) %>%
      dplyr::mutate(day = lubridate::day(`First Pollination Date`),
             month = lubridate::month(`First Pollination Date`),
             year = lubridate::year(`First Pollination Date`)) %>%
      .[complete.cases(.),]
    result = result %>% dplyr::filter(`First Pollination Date` >= input$dateRange1[1], `First Pollination Date` <= input$dateRange1[2])
    
    if((input$dateRange1[2] - input$dateRange1[1]) < 31){
      result = result %>%
        dplyr::filter(Location %in% input$site1,
               `First Pollination Date` >= input$dateRange1[1],
               `First Pollination Date` <= input$dateRange1[2]) %>%
        dplyr::group_by(day) %>% 
        dplyr::tally()
      grp = result$day
      group_name = "Daily"
    } else if((lubridate::year(input$dateRange1[2]) == lubridate::year(input$dateRange1[1]))){
      result <- dplyr::filter(result, Location %in% input$site1,
                       lubridate::year(`First Pollination Date`) == lubridate::year(input$dateRange1[1]),
                       lubridate::month(`First Pollination Date`) %between% c(lubridate::month(input$dateRange1[1]),lubridate::month(input$dateRange1[2]))) %>% 
        dplyr::group_by(month) %>% 
        dplyr::tally() 
      grp = month.abb[result$month]
      group_name = "Monthly"
    }  else if((lubridate::year(input$dateRange1[2]) != lubridate::year(input$dateRange1[1]))){ 
      result <- dplyr::filter(result, Location %in% input$site1, 
                       lubridate::year(`First Pollination Date`) %between% c(lubridate::year(input$dateRange1[1]), lubridate::year(input$dateRange1[2])))  %>% 
        dplyr::group_by(year) %>% dplyr::tally()
      
      grp = result$year
      group_name = "Yearly"
    }
    
   
    # PLOT POLLINATION
    
    highchart() %>%
      hc_add_series(data = result$n, type = "column",name = paste(group_name, " total crosses"),events = list(click = js_click_line)) %>%
      hc_xAxis( categories = grp,tickmarkPlacement="on") %>%
      hc_exporting(enabled = TRUE) %>% 
      hc_tooltip(crosshairs = TRUE, backgroundColor = "#FCFFC5",
                 shared = TRUE, borderWidth = 2) %>%
      hc_add_theme(hc_theme_elementary())

```


<br><br>


## Data Tables

Data tables has many useful options such as search and filters (just below the table headers) as describe below; 
<br>

*  Search Box: Located at top-right corner of data table. Use to search values within the table <br>
*  Table Filters: Filter within column. Multiple filters are posible for a fine view
*  Page Length: Shows number of entries to display in the table
*  Use the <b>Download</b> button at top-right corner of table to download data. Selecting specific rows, subsets the data downloaded.

<br>


```{r,echo=FALSE}
DT::datatable(head(banana), filter = "top")
```


<br><br>


## Acknowledgements

This dashboard and the report are made possible through the use of the R software and the following packages:

+ shiny: Winston Chang, Joe Cheng, JJ Allaire, Yihui Xie and Jonathan McPherson (2015). shiny: Web Application Framework for R. R package version 0.12.1. http://CRAN.R-project.org/package=shiny

+ DT:Yihui Xie (2014). DT: R Interface to the jQuery Plug-in DataTables. R package version 0.0.34. http://rstudio.github.io/DT

+ data.table: M Dowle, T Short, S Lianoglou, A Srinivasan with contributions from R Saporta and E Antonyan (2014). data.table: Extension of data.frame. R package version 1.9.4. http://CRAN.R-project.org/package=data.table

+ shinyBS: Eric Bailey (2015). shinyBS: Twitter Bootstrap Components for Shiny. R package version 0.61. http://CRAN.R-project.org/package=shinyBS

+ highcharter: Joshua Kunst, Nuno Agostinho (hchart.survfit, densities and hc_add_series_scatter), Eduardo Flores (2017). A Wrapper for the 'Highcharts' Library.  

+ collapsibleTree: A Khan, D Shah, M Bostock (2018). Interactive Collapsible Tree Diagrams using 'D3.js'. https://cran.r-project.org/web/packages/collapsibleTree/index.html

+ dplyr: Hadley Wickham, Romain Fran?ois, Lionel Henry, Kirill M?ller, RStudio. A Grammar of Data Manipulation. https://cran.r-project.org/web/packages/dplyr/index.html
<br> <br>
